---
title: "Introduction to geobr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic-Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The [**geobr**](https://github.com/ipeaGIT/geobr) package provides quick and easy access to shapefiles of the Brazilian Institute of Geography and Statistics (IBGE) and other official spatial data sets of Brazil. The syntax of all **geobr** functions operate a simple logic that allows users to easily download a wide variety of official spatial data sets with updated geometries, harmonized attributes and geographic projections across geographies and years. This vignette presents is a quick intro to **geobr**.


## Installation

The package will soon be available on CRAN. In the meantime, you can install it using the `devtools` package.

```{r }
library(devtools)
library(digest)

# devtools::install_github("ipeaGIT/geobr")
  library(geobr)

# Other libraries we'll use
  library(ggplot2)
  library(sf)
  library(dplyr)
  library(rio)

```



## Download shape files as `sf` objects

The territory of Brazil is organized in political-administrative and statistical areas at different levels: 
* States
* Regions
* Meso regions
* Micro regions
* Municipalities
* Census weighting areas
* Census tracts
* Statistical Grid 

The syntax of all *geobr* functions operate one the same logic, so the code to download and read those geographies is quite intuitive. Here are a few examples. Note the number of polygons and boundaries at each geographic level has changed over the years years, so it's important to specify the `year` of reference you want to download.


Read an specific geographic area at a given year
```{r message=FALSE,warning=FALSE,results='hide'}
# download data
  state <- read_state(code_state=11, year=2000)              # State
  micro <- read_micro_region(code_micro=110205, year=2000)   # Micro region
  munic <- read_municipality(code_muni=1200179, year=2017)   # Municipality

```


Read all geographic areas within a state at a given year
```{r message=FALSE,warning=FALSE,results='hide'}
# download data
  micro <- read_micro_region(code_micro=15, year=2013)       # Micro region
  munic <- read_municipality(code_muni= 33, year=2010)       # Municipality

# Or simply use the two-digit abbreviation of a state
  micro <- read_micro_region(code_micro="PA", year=2000)     # Micro region
  munic <- read_municipality(code_muni= "RJ", year=2010)     # Municipality

```


The user can also read all geographic areas in the country in a single line of code.
```{r message=FALSE,warning=FALSE,results='hide'}
# download data
  state <- read_state(code_state="all", year=2010)           # State
  micro <- read_micro_region(code_micro="all", year=2015)    # Micro region
  munic <- read_municipality(code_muni="all", year=2018)     # Municipality

```


## Plot the data


Once you've downloaded the data, it is really simple to plot mas using ggplot2`.


```{r , fig.height = 8, fig.width = 8, fig.align = "center",message=FALSE,warning=FALSE}

# No plot axis
  no_axis <- theme(axis.title=element_blank(),
                   axis.text=element_blank(),
                   axis.ticks=element_blank())



# Plot all Brazilian states
  ggplot() + 
    geom_sf(data=state, fill="#2D3E50", color="#FEBF57", size=.15, show.legend = FALSE) + 
    geom_sf_text(data=state, aes(label = abbrev_state), colour = "white") +
    labs(subtitle="States", size=8) + 
    theme_minimal() +
    no_axis
  
```

One can also download the meso regions of a particular state, such as the Amazonas state:

```{r , fig.height = 8, fig.width = 8, fig.align = "center",message=FALSE,warning=FALSE}
# Meso region download
  am_meso <- read_meso_region( code_meso = "AM", year= 2010)
  
# plot
  ggplot() +
    geom_sf(data=am_meso, fill="#2D3E50", color="#FEBF57", size=.15, show.legend = FALSE) +
    geom_sf_text(data=am_meso, aes(label = paste0(name_meso,"\n Code=",code_meso)), colour = "white",size = 2.6) +
    labs(subtitle="Meso regions in Amazonas states (code 13)", size=8) + 
    theme_minimal() +
    no_axis
  
```
Similarly, one can easily download all municipalities of particular regions:


```{r , fig.height = 8, fig.width = 8, fig.align = "center",message=FALSE,warning=FALSE}
# Download all municipalities
  all_muni <- read_municipality( code_muni = "all", year= 2010)
  
# Keep municipalities in the North and Central-West regions
  myregion <- subset(all_muni, name_region == "Norte")
    

# micro region plot
  ggplot() + 
    geom_sf(data=myregion, fill="#2D3E50", color="#FEBF57", size=.15, show.legend = FALSE) +
    labs(subtitle="Municipalities of the North region", size=8) + 
    theme_minimal() +
    no_axis
    
```


## Thematic maps

The next step is to combine  data from ***geobr*** package with other data sets to create thematic maps. In this example, we will be using data from the (Atlas of Human Development (a project of our colleagues at Ipea)[http://atlasbrasil.org.br/2013/pt/download/] to create a choropleth map showing the spatial variation of Life Expectancy at birth across Brazilian municipalities.

#### Merge external data

First, we need to download the Life Expectancy data set and merge it to our spatial database. The municipality code is our key to join these two databases.

```{r , fig.height = 8, fig.width = 8, fig.align = "center",message=FALSE,warning=FALSE}

# download Life Expectancy data
  adh <- rio::import("http://atlasbrasil.org.br/2013/data/rawData/atlas2013_dadosbrutos_pt.xlsx",
                     which = "MUN 91-00-10")

# keep only information for the year 2010 and the columns we want
  adh <- subset(adh, ANO == 2010)
  adh <- dplyr::select(adh, ANO,UF,Codmun7, IDHM, ESPVIDA)

# Download the sf of all Brazilian municipalities
  all_muni <- read_municipality( code_muni= "all", year= 2010)

  
# joind the databases
  all_muni <-left_join(all_muni, adh, by = c("code_muni" = "Codmun7"))
```

#### Plot thematic map

```{r , fig.height = 8, fig.width = 8, fig.align = "center",message=FALSE,warning=FALSE}

  ggplot() + 
    geom_sf(data=all_muni, aes(fill=ESPVIDA), color= NA, size=.15) + 
      labs(subtitle="Life Expectancy at birth, Brazilian Municipalities 2010", size=8) + 
      scale_fill_distiller(palette = "Blues", name="Life Expectancy", limits = c(65,80)) +
      theme_minimal() +
      no_axis

```

