---
title: "Introduction to geobr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic-Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The [**geobr**](https://github.com/ipeaGIT/geobr) package provides quick and easy access to official spatial data sets of Brazil. The syntax of all **geobr** functions operate on a simple logic that allows users to easily download a wide variety of data sets with updated geometries and harmonized attributes and geographic projections across geographies and years. This vignette presents a quick intro to **geobr**.


## Installation

You can install geobr from CRAN or the development version to use the latest features.

```{r eval=FALSE, message=FALSE, warning=FALSE}
# From CRAN
  install.packages("geobr")

# Development version
  utils::remove.packages('geobr')
  devtools::install_github("ipeaGIT/geobr", subdir = "r-package")

```

Now let's load the libraries we'll use in this vignette.

```{r message=FALSE, warning=FALSE, results='hide'}
  library(geobr)
  library(ggplot2)
  library(sf)
  library(dplyr)
  library(rio)
```



## General usage

### Available data sets

The geobr package covers 21 spatial data sets, including a variety of political-administrative and statistical areas used in Brazil. You can view what data sets are available using the `list_geobr()` function.

```{r message=FALSE, warning=FALSE}
# Available data sets
datasets <- list_geobr()

print(datasets, n=21)

```


## Download spatial data as `sf` objects

The syntax of all *geobr* functions operate one the same logic, so the code to download the data becomes intuitive for the user. Here are a few examples.

Download an specific geographic area at a given year
```{r eval=FALSE, message=FALSE, warning=FALSE, results='hide'}
# State of Sergige
  state <- read_state(code_state="SE", year=2018)

# Municipality of Sao Paulo
  muni <- read_municipality( code_muni = 3550308, year=2010 )

```


Download all geographic areas within a state at a given year
```{r eval=FALSE, message=FALSE, warning=FALSE, results='hide'}
# All municipalities in the state of Alagoas
  muni <- read_municipality(code_muni= "AL", year=2007)

# All census tracts in the state of Rio de Janeiro
  cntr <- read_census_tract(code_tract = "RJ", year = 2010)

```

If the paramter `code_` is not passed to the function, geobr returns the data for the whole country by default.

```{r message=FALSE, warning=FALSE, results='hide'}
meso <- read_intermediate_region(year=2017)
states <- read_state(year=2014)

```


## Importante note about data resolution

All functions to download polygon data such as states, municipalites etc. have a `tp` argument as in data ***type***. When `tp = "original"`, geobr will return the original data set with high resolution at detailed geographic scale (see documentation). By default, however, `tp = "simplified"` and geobr returns data set geometries with simplified borders to improve speed of downloading and plotting the data.




## Plot the data


Once you've downloaded the data, it is really simple to plot maps using `ggplot2`.


```{r message=FALSE, warning=FALSE, fig.height = 8, fig.width = 8, fig.align = "center"}
# Remove plot axis
  no_axis <- theme(axis.title=element_blank(),
                   axis.text=element_blank(),
                   axis.ticks=element_blank())



# Plot all Brazilian states
  ggplot() +
    geom_sf(data=states, fill="#2D3E50", color="#FEBF57", size=.15, show.legend = FALSE) +
    labs(subtitle="States", size=8) +
    theme_minimal() +
    no_axis

```

Plot all the municipalities of a particular state, such as Rio de Janeiro:

```{r message=FALSE, warning=FALSE, results='hide', fig.height = 8, fig.width = 8, fig.align = "center"}
# Download all municipalities of Rio
  all_muni <- read_municipality( code_muni = "RJ", year= 2000)
```

```{r message=FALSE, warning=FALSE, fig.height = 8, fig.width = 8, fig.align = "center"}

# plot
  ggplot() +
    geom_sf(data=all_muni, fill="#2D3E50", color="#FEBF57", size=.15, show.legend = FALSE) +
    labs(subtitle="Municipalities of Rio de Janeiro, 2000", size=8) +
    theme_minimal() +
    no_axis
```


## Thematic maps

The next step is to combine  data from ***geobr*** package with other data sets to create thematic maps. In this example, we will be using data from the (Atlas of Human Development (a project of our colleagues at Ipea))[http://atlasbrasil.org.br/2013/pt/download/] to create a choropleth map showing the spatial variation of Life Expectancy at birth across Brazilian states.

#### Merge external data

First, we need to download the Life Expectancy data set and merge it to our spatial database. The two-digit abbreviation of state name is our key column to join these two databases.

```{r message=FALSE, warning=FALSE, results='hide', fig.height = 8, fig.width = 8, fig.align = "center"}
# download Life Expectancy data
  adh <- rio::import("http://atlasbrasil.org.br/2013/data/rawData/Indicadores%20Atlas%20-%20RADAR%20IDHM.xlsx", which = "Dados")

# keep only information for the year 2010 and the columns we want
  adh <- subset(adh, ANO == 2014)

# Download the sf of all Brazilian states
  # states <- read_state(year= 2014)

# joind the databases
  states <-left_join(states, adh, by = c("abbrev_state" = "NOME_AGREGA"))
```

#### Plot thematic map

```{r message=FALSE, warning=FALSE, fig.height = 8, fig.width = 8, fig.align = "center" }
  ggplot() +
    geom_sf(data=states, aes(fill=ESPVIDA), color= NA, size=.15) +
      labs(subtitle="Life Expectancy at birth, Brazilian States, 2014", size=8) +
      scale_fill_distiller(palette = "Blues", name="Life Expectancy", limits = c(65,80)) +
      theme_minimal() +
      no_axis

```

