---
title: "Introduction to geobr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic-Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The [**geobr**](https://github.com/ipeaGIT/geobr) package provides quick and easy access to shapefiles of the Brazilian Institute of Geography and Statistics (IBGE) and other official spatial data sets of Brazil. The syntax of all **geobr** functions operate a simple logic that allows users to easily download a wide variety of official spatial data sets with updated geometries, harmonized attributes and geographic projections across geographies and years. This vignette presents is a quick intro to **geobr**.


## Installation

The package will soon be available on CRAN. In the meantime, you can install it using the `devtools` package.

```{r message=FALSE, warning=FALSE}
# Load geobr and other libraries we'll use
  library(geobr)
  library(ggplot2)
  library(sf)
  library(dplyr)
  library(rio)

```



## Download shape files as `sf` objects

The territory of Brazil is organized in political-administrative and statistical areas at different levels: 
* Country
* States
* Regions
* Meso regions
* Micro regions
* Municipalities
* Census weighting areas
* Census tracts
* Statistical Grid 

The syntax of all *geobr* functions operate one the same logic, so the code to download and read those geographies is quite intuitive. Here are a few examples. Note the number of polygons and boundaries at each geographic level has changed over the years years, so it's important to specify the `year` of reference you want to download.


Read an specific geographic area at a given year
```{r message=FALSE, warning=FALSE, results='hide'}
# download data
  state <- read_state(code_state="SE", year=2018)          # State
  micro <- read_micro_region(code_micro=160101, year=2000) # Micro region

```


Read all geographic areas within a state at a given year
```{r message=FALSE, warning=FALSE, results='hide'}
# download data
  meso <- read_meso_region(code_meso="PA", year=2018)      # Meso region
  muni <- read_municipality(code_muni= "AL", year=2007)    # Municipality

```


The user can also read all geographic areas in the country in a single line of code.
```{r message=FALSE, warning=FALSE, results='hide'}
# download data
  state <- read_state(code_state="all", year=1991)         # State
  meso <- read_meso_region(code_meso="all", year=2001)     # Meso region

```


## Plot the data


Once you've downloaded the data, it is really simple to plot maps using `ggplot2`.


```{r message=FALSE, warning=FALSE, fig.height = 8, fig.width = 8, fig.align = "center"}

# No plot axis
  no_axis <- theme(axis.title=element_blank(),
                   axis.text=element_blank(),
                   axis.ticks=element_blank())



# Plot all Brazilian states
  ggplot() + 
    geom_sf(data=state, fill="#2D3E50", color="#FEBF57", size=.15, show.legend = FALSE) + 
    labs(subtitle="States", size=8) + 
    theme_minimal() +
    no_axis
  
```

All the municipalities of a particular state, such as Rio de Janeiro:

```{r message=FALSE, warning=FALSE, results='hide', fig.height = 8, fig.width = 8, fig.align = "center"}
# Download all municipalities of Rio
  all_muni <- read_municipality( code_muni = "RJ", year= 2000)
```

```{r message=FALSE, warning=FALSE, fig.height = 8, fig.width = 8, fig.align = "center"}

# plot
  ggplot() + 
    geom_sf(data=all_muni, fill="#2D3E50", color="#FEBF57", size=.15, show.legend = FALSE) +
    labs(subtitle="Municipalities of Rio de Janeiro, 2000", size=8) + 
    theme_minimal() +
    no_axis
```


## Thematic maps

The next step is to combine  data from ***geobr*** package with other data sets to create thematic maps. In this example, we will be using data from the (Atlas of Human Development (a project of our colleagues at Ipea)[http://atlasbrasil.org.br/2013/pt/download/] to create a choropleth map showing the spatial variation of Life Expectancy at birth across Brazilian states.

#### Merge external data

First, we need to download the Life Expectancy data set and merge it to our spatial database. The two-digit abbreviation of state names is our key to join these two databases.

```{r message=FALSE, warning=FALSE, results='hide', fig.height = 8, fig.width = 8, fig.align = "center"}

# download Life Expectancy data
  adh <- rio::import("http://atlasbrasil.org.br/2013/data/rawData/Indicadores%20Atlas%20-%20RADAR%20IDHM.xlsx", which = "Dados")

# keep only information for the year 2010 and the columns we want
  adh <- subset(adh, ANO == 2014)
  
# Download the sf of all Brazilian states
  all_states <- read_state(code_state= "all", year= 2014)
  
# joind the databases
  all_states <-left_join(all_states, adh, by = c("abbrev_state" = "NOME_AGREGA"))
```

#### Plot thematic map

```{r message=FALSE, warning=FALSE, fig.height = 8, fig.width = 8, fig.align = "center" }
  ggplot() + 
    geom_sf(data=all_states, aes(fill=ESPVIDA), color= NA, size=.15) + 
      labs(subtitle="Life Expectancy at birth, Brazilian States, 2014", size=8) + 
      scale_fill_distiller(palette = "Blues", name="Life Expectancy", limits = c(65,80)) +
      theme_minimal() +
      no_axis

```

