---
title: "Introduction to geobr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic-Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The [**geobr**](https://github.com/ipeaGIT/geobr) package provides quick and easy access to official spatial data sets of Brazil. The syntax of all **geobr** functions operate on a simple logic that allows users to easily download a wide variety of data sets with updated geometries and harmonized attributes and geographic projections across geographies and years. This vignette presents a quick intro to **geobr**.


## Installation

You can install geobr from CRAN or the development version to use the latest features.

# {.tabset}

## R
```{r eval=FALSE, message=FALSE, warning=FALSE}
# From CRAN
  install.packages("geobr")

# Development version
  utils::remove.packages('geobr')
  devtools::install_github("ipeaGIT/geobr", subdir = "r-package")

```

## Python

```{r eval=FALSE, message=FALSE, warning=FALSE}

import geobr
import matplotlib.pyplot as plt
import pandas as pd

%matplotlib inline
```

teste 2 

## R
```{r eval=FALSE, message=FALSE, warning=FALSE}
# From CRAN
  install.packages("geobr")

# Development version
  utils::remove.packages('geobr')
  devtools::install_github("ipeaGIT/geobr", subdir = "r-package")

```

## Python
```{r eval=FALSE, message=FALSE, warning=FALSE}

import geobr
import matplotlib.pyplot as plt
import pandas as pd

%matplotlib inline
```
